---
title: "IDEEA"
output: rmarkdown::html_vignette
params:
  solve: false
vignette: >
  %\VignetteIndexEntry{IDEEA}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = FALSE,
  comment = "#>"
)
```

# Configuration
```{r, eval = T}
# number of regions
nreg <- 5 
# nreg <- 32 
```

```{r setup, message=FALSE, warning=FALSE}
library(IDEEA)
library(tidyverse)
library(data.table)
library(sf)
library(glue)

regN <- glue("reg", nreg)
regN_off <- paste0(regN, "_off")

```

## Regions

```{r regions}
ideea_sf <- get_ideea_map(nreg, offshore = F, reg_off = T)
plot(ideea_sf[regN], key.width = lcm(4))

```

## Electricity model
```{r}
# Import electricity model objects
elc <- ideea_modules$electricity[[regN]]
names(elc)
# summary(elc)

# A model with the default configuration
mod <- newModel(
  name = glue("IDEEA_ELC_{regN}"),
  desc = "IDEEA electricity example model",
  region = unique(ideea_sf[[regN]]),
  discount = 0.05,
  calendar = ideea_modules$calendars$calendar_d365_h24,
  horizon = ideea_modules$calendars$horizon_2020_2060_by_10,
  data = newRepository(
    name = glue("repo_electricity_{regN}"),
    elc$repo_comm, # declaration of commodities used in the model
    # supply & import
    elc$repo_supply, # primary energy supply (markets)
    elc$repo_geoccs, # CCS storage potential by region
    # Generating technologies
    elc$ECOA, # coal-fired generators
    elc$EGAS, # gas-fired generators
    elc$ENUC, # nuclear generators
    elc$EHYD, # hydro generators
    elc$ESOL_w_const, # solar generators with constraints on capacity
    # elc$ESOL_w_land, # solar generators with constraints on land
    elc$EWIN_w_const, # wind generators with constraints on capacity
    # elc$EWIN_w_land, # wind generators with constraints on land
    elc$EWIF_w_const, # wind generators with constraints on capacity
    # elc$EWIF_w_land, # wind generators with constraints on used surface area
    elc$EBIO, # biomass generators
    elc$repo_ccstechs, # CCS technologies
    # capacity (weather) factors
    elc$WSOL, # solar capacity factor
    elc$WWIN, # onshore wind capacity 
    elc$WWIF, # offshore wind capacity 
    elc$WHYD, # hydro capacity factor
    # battery
    elc$STG_BTR, # battery storage with yearly cycle
    # STG_BTR_daily = STG_BTR_daily, # battery storage with daily cycle
    # transmission
    elc$repo_transmission_ac, # HVAC transmission network
    elc$repo_transmission_dc, # HVDC transmission network with converters
    # assumptions / policies
    elc$NO_NEW_HYDRO, # no new hydro capacity
    elc$NO_NEW_NUCLEAR, # no new nuclear capacity
    # elc$NO_NEW_WIND, # no new wind capacity
    # elc$NO_NEW_SOLAR, # no new solar capacity
    # demand
    elc$DEMELC_BY, # electricity demand in the base year
    elc$DEMELC_2X # electricity demand growth
  )
)

```

## Scenarios
Let's consider two scenarios for the electricity sector in 2050. The first scenario is a reference scenario (**REF**) without any policy intervention. The second scenario is a scenario with a carbon emissions constraint (**CAP**), targeting zero emissions by 2060 (linear reduction from 2030).  

Here we solve both scenarios for 2050 with full 8760 hours resolution and a subset of 1 day per month, 288 hours (24 * 12) in total. The subset scenario is relatively fast to solve and can be used for drafting and testing purposes. Though it does not provide a full picture of intermittent renewable generation and can downplay the role of storage and transmission.


### Reference scenario
The **REF** scenario is a business-as-usual that includes all the existing technologies that are already in the model. 

#### Solving for 288 hours
The 288 hours scenario can be solved with open source solvers like HiGHS or Cbc in minutes. Here we use the HiGHS solver with the barrier method and Julia interface.
```{r}
# (optional) set default solver
set_default_solver(solver_options$julia_highs_barrier)
set_progress_bar(type = "cli") # output progress bar in the console

# Sub-setting and interpolating of the model parameters
scen_REF_sub <- interpolate_model(
  mod,
  name = "REF_sub", # indicating 'subset'
  calendar = ideea_modules$calendars$calendar_d365_h24_subset_1day_per_month,
  horizon = ideea_modules$horizons$horizon_2050
)

# Writing the model script and data
scen_REF_sub <- write_sc(
  scen_REF_sub,
  solver = solver_options$julia_highs_barrier
)
scen_REF_sub@status # info
scen_REF_sub@misc$tmp.dir # temporary directory

# Solving the scenario
scen_REF_sub <- solve_scenario(scen_REF_sub, wait = T)

# Reading the results
scen_REF_sub <- read(scen_REF_sub)
summary(scen_REF_sub)

# (optional) saving the scenario on disk
scen_REF_sub <- save_scenario(scen_REF_sub)

```

#### Solving for 8760 hours
The full-year scenario requires more computational resources and time. The generation of the problem also varies by the software. Here we use Python/Pyomo and CPLEX solver. The one-year scenario can be solved within an hour on a consumer-grade desktop or laptop. Barrier algorithm benefits from parallel processing and is normally faster on multi-core CPUs.

```{r}
# Interpolating the model parameters
scen_REF <- interpolate(
  mod,
  name = "REF",
  # calendar = ideea_modules$calendars$calendar_d365_h24_full, # already set in the model
  horizon = ideea_modules$horizons$horizon_2050
)

# Writing the model script and data
scen_REF <- write_sc(
  scen_REF,
  solver_options$pyomo_cplex_barrier
  # solver_options$julia_cplex_barrier
  # solver = solver_options$gams_gdx_cplex_barrier
  # solver = solver_options$gams_gdx_cplex_parallel
  # solver = solver_options$julia_highs_barrier
  # solver = solver_options$julia_highs_parallel
)

# Solving the scenario
scen_REF <- solve_scenario(scen_REF, wait = T)

# Reading the results
scen_REF <- read(scen_REF)
summary(scen_REF)

# (optional) saving the scenario on disk
scen_REF <- save_scenario(scen_REF)

```

### Carbon emissions constraint scenario
The **CAP** scenario is a scenario with a carbon emissions constraint. Here we use pre-defined CO2 emissions constraint on the national level, assuming emissions trading. The constraint is saved in `ideea_modules[[regN]]$CO2_CAP` object.

#### Solving for 288 hours
```{r}
# Policy constraint 
CO2_CAP <- ideea_modules$electricity[[regN]]$CO2_CAP

# Sub-setting and interpolating of the model parameters
scen_CAP_sub <- interpolate_model(
  mod,
  name = "CAP_sub", # indicating 'subset'
  CO2_CAP, # carbon emissions constraint
  calendar = ideea_modules$calendars$calendar_d365_h24_subset_1day_per_month,
  horizon = ideea_modules$horizons$horizon_2050
)

# Writing the model script and data
scen_CAP_sub <- write_sc(
  scen_CAP_sub,
  # solver = solver_options$gams_gdx_cplex_parallel
  solver = solver_options$julia_highs_barrier
)

# Solving the scenario
scen_CAP_sub <- solve_scenario(scen_CAP_sub, wait = T)

# Reading the results
scen_CAP_sub <- read(scen_CAP_sub)
summary(scen_CAP_sub)

# (optional) saving the scenario on disk
scen_CAP_sub <- save_scenario(scen_CAP_sub)

```


#### Solving for 8760 hours 
Similarly, the full-year scenario can be solved with the same solver and settings as **REF**.
```{r}
# Interpolating the model parameters
scen_CAP <- interpolate(
  mod,
  name = "CAP",
  CO2_CAP, # carbon emissions constraint
  horizon = ideea_modules$horizons$horizon_2050
)

# Writing the model script and data
scen_CAP <- write_sc(
  scen_CAP,
  solver_options$pyomo_cplex_barrier
  # solver_options$julia_cplex_barrier
  # solver = solver_options$gams_gdx_cplex_barrier
  # solver = solver_options$gams_gdx_cplex_parallel
  # solver = solver_options$julia_highs_barrier
  # solver = solver_options$julia_highs_parallel
)

# Solving the scenario
scen_CAP <- solve_scenario(scen_CAP, wait = T)

# Reading the results
scen_CAP <- read(scen_CAP)
summary(scen_CAP)
isInMemory(scen_CAP) 

# (optional) saving the scenario on disk
scen_CAP <- save_scenario(scen_CAP)
isInMemory(scen_CAP) # FALSE means the scenario is saved on disk
summary(scen_CAP) # check RAM size 

```


## Results
... an example of processing some results ...
```{r}
# Combine scenarios in a named or unnamed list
# (in the case of unnamed list, scenario names are used)
sns <- list(scen_REF_sub, scen_CAP_sub)
# sns <- list(scen_REF_sub, scen_REF, scen_CAP_sub, scen_CAP)

```

Use `getData()` function to retrieve variables and parameters from solved scenario objects. See `?getData` for more details.
```{r}
# Objective function value (total discounted cost)
getData(sns, "vObjective", merge = T)
```
The "subset" time-slices scenarios normally show lower objective. 

Compare technological capacity by scenario.
```{r}
vTechCap <- getData(sns, "vTechCap", merge = T, process = T, tech_ = "^E",
                    digits = 2, drop.zeros = T) |> 
  drop_process_cluster() |> # removes cluster indicators from process names
  drop_process_vintage() |> # removes vintage indicators from process names
  group_by(scenario, process, year) |> # group and aggregate for figures/tables
  summarise(GW = sum(value), .groups = "drop") |>
  filter(!grepl("EDC", process)) |>
  as.data.table()
vTechCap |> pivot_wider(names_from = process, values_from = GW)

ggplot(vTechCap) +
  geom_bar(aes(x = scenario, y = GW, fill = process), stat = "identity") +
  scale_fill_viridis_d(option = "H", direction = -1) +
  theme_bw() 

```


### Reports
```{r, eval=params$solve}
# Collect all time-slices from a scenario
all_slices <- scen_CAP@settings@calendar@timetable$slice
# Plot "snapshot" of an system hourly operation for a particular day/year
ideea_snapshot(scen_CAP, YEAR = 2055, SLICE = all_slices[grepl("d015", all_slices)])

ideea_snapshot(scen_CAP, YEAR = 2060, 
               SLICE = all_slices[grepl("d015", all_slices)],
               return_data = F) +
  facet_grid(region~year, scales = "free_y")

```

tbc...

