---
title: "IDEEA: regions and maps"
output:
  html_document:
    df_print: paged
  pdf_document: default
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = TRUE,
  fig.width = 6,
  fig.height = 6,
  comment = "#>"
)
```

Geographical information system (GIS) files, or maps are used in the project for identification of locations of solar and wind energy potential and visualisation of the model data and modeling results. IDEEA package has embedded maps with alternative regional divisions, starting from one region (the country level) to 37 regions (each state as a region) or 46 zones where few regions are split into several distant territories (such as Puduchery). The maps are available in *simple feature* (`sf`) format and can be called using the `get_ideea_maps` function:
```{r, message=FALSE, warning=FALSE}
library(IDEEA)
ideea_sf <- get_ideea_map()
plot(ideea_sf["region"], main = paste0(length(unique(ideea_sf$region)), 
                                       " regions map with offshores"))
```

The function returns `sf` object with region short (`region` or `reg{N}`) and long names (`name`), names with offshore areas (`reg_off`), as well as further information (see `?get_ideea_map` for details).  

```{r}
get_ideea_map() # returns 5 regions by default
```
Administratively, India is organized into 28 states and 8 Union territories, initially restructured according to linguistic lines as per The States Reorganization Act 1956. In a specific modelling framework that allows for regional modification, a GIS map initially divides the country into 46 regions, incorporating all states and Union territories with certain modifications. This division counts Daman and Diu as separate entities and treats each district of Puducherry as its own region, reflecting its composition of formerly separate French colonies. For simplification, alternative regional models consolidate these areas differently, reducing the number of regions to 36, 34, 32, 7, and 5, by merging split areas (such as Puducherry) into a nearby entity and combining Daman and Diu. This approach provides examples of alternative levels of details for modeling India's energy system, where 32-37 aggregation options are closest to the administrative boundaries, but has splits and aggregates of some regions with no political context. Scholars can bring to use their own shape-files and regional divisions for their studies, publications, and consulting. 


```{r}
ideea_sf <- get_ideea_map(nreg = 36, offshore = TRUE, islands = TRUE)
# ideea_sf
plot(ideea_sf["region"], main = paste0(length(unique(ideea_sf$region)), " regions map"))
```

The IDEEA map can me acquired for 1, 5, 7, 32, 34, 36, 37 or 46 regions. The `offshore` and `islands` arguments allow to include or exclude offshore areas and islands, respectively. The `rename` argument allows to rename the columns to `region` and `name` for the region short and long names. The `reg_off` argument allows to add a column with offshore areas to the region names. The `aggregate` argument allows to aggregate geometries of regions with the same name. 

```{r}
ideea_sf <- get_ideea_map(nreg = 1)
plot(ideea_sf["region"], main = paste0(length(unique(ideea_sf$region)), " regions map"))
```

The mapping of alternative regional divisions can be found in the `ideea_data$reg_tbl` data frame where every aggregation is represented in a separate colun with names `reg{N}` and `name{N}`. The `offshore` and `mainland` columns allow to distinguish between offshore and mainland areas. 

```{r}
ideea_data$reg_tbl |> head()
```

Alternative aggregations can be added using this table and the `sf` object `get_ideea_map(aggregate = F, rename = FALSE, islands = T)`.  

