---
title: "Reports"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Analysis of the model output

An example of a report that can be generated from the model output, using results from the IDEEA electricity model.


```{r setup, warning=FALSE, message=FALSE}
library(IDEEA)
library(tidyverse)
```


### Load the model output
Here we use pre-saved results from the IDEEA electricity model, 5 or 32 regions, as described in the `vignette("electricity")`.

```{r}
# load reference scenario
load_scenario(file.path(ideea_scenarios(), "REF_IDEEA_ELC_reg32_d365_h24_2060"))
# load carbon constrained scenario
load_scenario(file.path(ideea_scenarios(), "CAP_IDEEA_ELC_reg32_d365_h24_2060"))

# combine scenarios in one list for comparison
sns <- list(
  REF_2060 = .scen$REF, 
  CAP_2060 = .scen$CAP
  )

```

<details><summary>Details</summary>
```{r}
# quick check
summary(.scen$REF)
summary(.scen$CAP)

summary(sns) # list with presolved IDEEA scenarios
```
</details>

### Compare the results

#### Objective 
Objective value of the model (minimized cost)
```{r}
getData(sns, "vObjective", merge = TRUE)
```

#### Installed capacity
<details><code></summary>
```{r}
vTechCap <- getData(
  scen = sns, 
  name = c("vTechCap"), # capacity variable of technologies
  tech_ = "^E", # select only electricity generators (start with "E")
  process = TRUE, # use unified set-name for all processes (rename 'tech' -> 'process')
  digits = 2, # round to 2 digits
  drop.zeros = T, # drop zero values
  merge = TRUE # merge tables
  ) |>
  drop_process_cluster() |> # erase cluster from process name
  drop_process_vintage() # erase vintage-year from process name

vStorageCap <- getData(
  scen = sns, 
  name = c("vStorageCap"), # capacity variables of storage technologies
  # tech_ = "^S", # select only storage technologies (start with "S")
  process = TRUE, # use unified set-name for all processes (rename 'tech' -> 'process')
  digits = 2, # round to 2 digits
  drop.zeros = T, # drop zero values
  merge = TRUE # merge tables
  ) |> 
  drop_process_vintage() # erase vintage-year from process name

# combine technologies and storage
vCap <- bind_rows(vTechCap, vStorageCap)

vCap$process |> unique()

```
</details>

```{r, echo=FALSE}
vCap |>
  group_by(scenario, process, year) |>
  summarise(GW = sum(value), .groups = "drop") |>
  pivot_wider(names_from = scenario, values_from = GW) |>
kableExtra::kable(
  caption = "Installed capacity of electricity generation and storage technologies, GW", 
  align = "c"
)
```


#### Transmission capacity
```{r}
vTradeCap <- getData(sns, "vTradeCap", merge = TRUE, process = T)
  
  
```


#### Generation mix
```{r, fig.cap="Electricity generation mix by technology, TWh"}
vTechOut_ELC <- getData(sns, "vTechOut", tech_ = "^E", process = T, digits = 2,
                        drop.zeros = TRUE, merge = TRUE) |>
  drop_process_cluster() |>
  drop_process_vintage()

vTechOut_ELC_year <- vTechOut_ELC |>
  group_by(scenario, process, year) |>
  summarise(TWh = sum(value) / 1e3, .groups = "drop") |>
  filter(!process %in% c("ELC2EDC", "EDC2ELC")) # drop converters

ggplot(vTechOut_ELC_year) +
  geom_bar(aes(x = scenario, y = TWh, fill = process), stat = "identity") +
  scale_fill_viridis_d(option = "H", direction = -1) +
  theme_bw()

```

#### Intraday generation profile




#### Emissions
```{r, echo=FALSE}
# Emissions from fuel combustion
# vEmsFuelTot <- getData(sns, "vEmsFuelTot", merge = TRUE)

# Emissions from processes (auxiliary commodities)
all_emis <- getData(sns, 
                    name = c("vTechAOut", "vEmsFuelTot"), 
                    process = T, digits = 2,
                    drop.zeros = TRUE, merge = TRUE)

all_emis$comm |> unique()

```


```{r, tab.cap="Emissions from fuel combustion and processes, CO2 Mt, kt others"}
all_emis |>
  group_by(scenario, comm, year) |>
  summarise(Mt = sum(value) / 1e3, .groups = "drop") |>
  pivot_wider(names_from = scenario, values_from = Mt) |>
  kableExtra::kable(
    # caption = "Emissions from fuel combustion and processes, CO2 Mt, kt others", 
    align = "c"
  )

```


