---
title: "IDEEA: transportation module"
output: pdf_document
# output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{transport}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.height = 5,
  fig.width = 7,
  comment = "#>"
)
```

```{r setup, warning=FALSE, message=FALSE}
library(IDEEA)
library(data.table)
```

Here we develop a transportation sector model for India with several sub-sectors or branches:  
* road transport  
  - light duty vehicles (LDV)  
  - medium duty vehicles (MDV)
  - heavy duty vehicles (HDV)  
  - buses   
    + city   
    + inter-city   
* rail transport  
* electric city transport  
  - trams and trolleys  
  - metro and short-distance trains  
* air transport  
  - domestic and international   
* water transport  
  - domestic  
  - international  

All transportation services deliver two types of services: passenger transportation, and freight. 


## Road transport



## Light Duty Vehicles (LDV)  

LDV or cars are modeled to deliver two types of transportation services, based on efficiency/fuel consumption per trip:  
* highway passenger*km (`P-LDV-HWY`)   
* city passenger*km (`P-LDV-CTY`) 


### Commodities
```{r}
GSL <- newCommodity(
  name = "GSL", 
  unit = "PJ", 
  timeframe = "ANNUAL")
```

### Transport services
```{r}
PLDVHWY <- newCommodity(
  name = "PLDVHWY", 
  desc = "LDV passengers transportation on highway",
  unit = "MPKm", # "million passenger * km", 
  timeframe = "YDAY")

PLDVCTY <- newCommodity(
  name = "PLDVCTY", 
  desc = "LDV passengers transportation in city",
  unit = "MPKm", # "million passenger * km", 
  timeframe = "YDAY")
```

### Infrastructure
```{r}
GSTATION <- newCommodity(
  name = "GSTATION", 
  desc = "Gas station",
  unit = "PJ", # number of cars to fuel a year
  timeframe = "ANNUAL")

CHARGEST <- newCommodity(
  name = "CHARGEST", 
  desc = "Electric car charing station",
  unit = "GWh", # number of cars to fuel a year
  timeframe = "ANNUAL")

```


### Gasoline cars
```{r}
# Technologies
LDVG <- newTechnology(
  name = "LDVG",
  desc = "Gasoline Light Duty Vehicles",
  input = data.frame(
    comm = c("GSL", "BIO"),
    unit = "PJ", # for simplicity and electrification scenarios, normally "PJ"
    group = "i"
  ),
  output = data.frame(
    comm = c("PLDVHWY",  # passenger * km, highway
             "PLDVCTY"), # passenger * km, city
    unit = "MPKm", # million passenger * km
    group = "o" # a number or string, "o" = output
  ),
  units = list(
    capacity = "1000 Vehicles",
    activity = "million km, city", # measured in one of the output commodities
    costs = "MUSD"
  ),
  cap2act = 10, # 10000 km per car a year => 10000 * 1000 / 1e6
  ceff = data.frame(
    comm = c("GSL", "BIO", "PLDVHWY", "PLDVCTY"),
    use2cact = c(NA, NA, 3, 2), # !!! calculate million km per unit of input
    share.up = c(NA, .1, .4, .8),
    cact2cout = c(NA, NA, 2, 3) # passengers per vehicles (assumption)
  ),
  olife = list(olife = 10),
  capacity = list(
    region = NA,
    year = 2022,
    stock = 500
  ),
  invcost = list(
    invcost = 15 # 15000 [USD/car] * 1000 [cars] / 1e6 => MUSD/1000 cars
  ),
  fixom = list(
    fixom = .5 # repairs a year, taxes and fees, tires, oil change, parking etc.
  ),
  # varom = list(
  #   varom = ... # could be "tires, oil change" 
  # )
  aux = list(
    acomm = c("GSTATION", "PM"), 
    unit = c("PJ", "kt")
  ), # N cars fueled a year
  aeff = list(
    acomm = c("GSTATION", "PM"),
    cap2ainp = c(1, NA), # estimate number of trips to gas-station a year
    act2aout = c(NA, .01) # find emissions data per km
    )
  )
  
draw(LDVG)

```

### Diesel cars
```{r}

```

### Hybrid cars
```{r}

```

### Plugged in hybrid cars
```{r}

```

### Electric cars
```{r}

```

### Hydrogen cars
```{r}

```



